Generalized universal instability: Transient linear amplification and subcritical turbulence

### 2015-02-26 replyToReferee.txt

We thank the referee for taking the time to carefully read the
manuscript and offer constructive feedback.

Regarding suggestion #1, discussing the formal mathematical
relationship between sensitivity of eigenmodes and transient growth
using concepts of pseudospectra:

We have added a paragraph discussing this subject to the end of
Section III.  However, we are reluctant to assert too strong a
connection to the mathematical theory of pseudospectra due to a
technical complication.  In the 'destabilizing perturbations' section
of Trefethen to which you refer, eq (13), the linear operator is given
a perturbation of norm bounded by some finite epsilon, causing the
eigenvalues to move by an amount that continuously goes to 0 as
epsilon -> 0. However, all evidence suggests that when shear is
changed from 0 to any finite value in the universal instability,
eigenvalues change discontinuously, with the change not going to 0
continuously as Ls/Ln -> infinity. In particular, the eigenvalues
discontinuously change from unstable to stable for a range of k_y, as
discussed in Refs. [25]-[26]. The reason for this discontinuity is
presumably the fact that the z domain must be infinite to resolve
eigenmode structures as shear goes to zero.  For any nonzero shear,
there is always some z for which the argument of J_0 is changed
significantly compared to the zero-shear case. Thus, to analyze the
role of magnetic shear in the universal instability using the formal
theory of pseudospectra, a level of mathematical care is required that
is beyond the scope of the present manuscript.

All of the other recommendations of the referee have been implemented,
and can be identified in the version of the new manuscript with
changes in red. For each item in the referee report, the relevant
locations of the manuscript changes are the following:

### 2015-02-26 universalModeTurbulencePaper.tex
Now responding to referee's comment 1

\author{Matt Landreman}
\email[]{mattland@umd.edu}
\affiliation{Institute for Research in Electronics and Applied Physics, University of Maryland, College Park, MD, 20742, USA}
\author{Gabriel G. Plunk}
\affiliation{Max Planck Institute for Plasma Physics, EURATOM Association and Max-Planck/Princeton Research Center for Plasma Physics}
\author{William Dorland}
\affiliation{Institute for Research in Electronics and Applied Physics, University of Maryland, College Park, MD, 20742, USA}
%\homepage[]{Your web page}
%\thanks{}
%\author{co-authors}

@@

\changed{
To demonstrate this principle across a range of parameters,
figure  \ref{fig:comparingTransientTo0Shear} demonstrates that the
correspondence between the finite-shear transients and zero-shear eigenmodes
is maintained throughout the evolution of the transient and for other
values of magetic shear.}
%This comparison is shown in figure \ref{fig:comparingTransientTo0Shear}.
In this figure, we display 25 independent simulations of linear transients,
varying $L_s/L_n$ between 20 and 300, and plotting the evolution of $\gamma = (2W(t))^{-1} dW/dt$ from the time
of peak growth rate to the time at which the energy begins to decay.
For all simulations shown, $k_y \rho_i = 2$ and $\nuee = 0.05 v_i/L_n$.
In each case, there is very close agreement between the instantaneous growth rate of the transient
and the `predicted' growth rate from the zero-shear dispersion relation.
Thus, we can conclude that the transiently growing structures are closely related to the zero-shear
eigenmodes.  
%Magnetic shear causes $k_{\perp}$ and $k_{||}$ to vary with time as the structure
%evolves (unlike the zero-shear case in which $k_{\perp}$ and $k_{||}$ are constants.)

@@

>>>>

\author{Matt Landreman}
\email[]{mattland@umd.edu}
\affiliation{Institute for Research in Electronics and Applied Physics, University of Maryland, College Park, MD, 20742, USA}
\author{Gabriel G. Plunk}
\affiliation{Max Planck Institute for Plasma Physics, Max-Planck/Princeton Research Center for Plasma Physics}
\author{William Dorland}
\affiliation{Institute for Research in Electronics and Applied Physics, University of Maryland, College Park, MD, 20742, USA}
%\homepage[]{Your web page}
%\thanks{}
%\author{co-authors}

@@

\changed{
To demonstrate this principle across a range of parameters,
figure  \ref{fig:comparingTransientTo0Shear} demonstrates that the
correspondence between the finite-shear transients and zero-shear eigenmodes
is maintained throughout the evolution of the transient, for a range of
values of magetic shear.}
%This comparison is shown in figure \ref{fig:comparingTransientTo0Shear}.
In this figure, we display 25 independent simulations of linear transients,
varying $L_s/L_n$ between 20 and 300, and plotting the evolution of $\gamma = (2W(t))^{-1} dW/dt$ from the time
of peak growth rate to the time at which the energy begins to decay.
For all simulations shown, $k_y \rho_i = 2$ and $\nuee = 0.05 v_i/L_n$.
In each case, there is very close agreement between the instantaneous growth rate of the transient
and the `predicted' growth rate from the zero-shear dispersion relation.
Thus, we can conclude that the transiently growing structures are closely related to the zero-shear
eigenmodes.  
%Magnetic shear causes $k_{\perp}$ and $k_{||}$ to vary with time as the structure
%evolves (unlike the zero-shear case in which $k_{\perp}$ and $k_{||}$ are constants.)

@@

\changed{In the context of sheared flows in neutral fluids, a
  relationship between transient growth and eigenvalue sensitivity has
  been formally proved using the notion of a pseudospectrum
  \cite{ReddyHenningson,TrefethenSubcritical,TrefethenEmbree}. The
  $\epsilon$-pseudospectrum of a linear operator can be defined as the
  union of the sets of eigenvalues obtained when the operator is
  perturbed by any other operator of norm $\le \epsilon$, providing a
  precise notion of the sensitivity of eigenvalues.  One can
  rigorously bound the amount of possible transient growth by the
  extent of the pseudospectrum into the unstable half of the complex
  plane \cite{ReddyHenningson} (eq (12) of
  \cite{TrefethenSubcritical}.)  However, for the universal
  instability, it is not clear that changing the magnetic shear from 0
  to a small finite value has the form of the operator perturbation
  used to define the pseudospectrum.  The reason is that the $z$
  domain is infinite, and for any nonzero shear, there is always some
  $z$ for which the argument of $J_0$ is changed significantly
  compared to the zero-shear case.  Indeed, when the shear changes
  from 0 to any small finite value, the eigenvalues (for a range of
  $k_y$) appear to change from unstable to stable discontinuously
  rather than continuously \cite{Ross,Tsang}.  Thus, it may or may not
  be possible to analyze the role of magnetic shear in the universal
  instability using the formal theory of pseudospectra, and we leave
  this careful mathematical analysis for future work.  }
